---
title: "Finnmaid pCO2"
author: "Jens Daniel Mueller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
workflowr:
  suppress_report: TRUE
---

# Data base

This analysis is based on pCO~2~ and SST observations from VOS Finnmaid, averaged for each crossing of previously defined sub-regions of the Baltic Sea (see [Schneider & Müller, 2018](https://www.springer.com/de/book/9783319616988)).
The sub-region "7. HGF" in front of Helsinki is selected, because this sub-region is usually crossed in the morning (before arrival in Helsinki) and again in the evening of the same day (after departure from Helsinki), therefore allowing to investigate daytime variability. Mean CT values for each visit are calculate from measured pCO~2~, SST, and a fixed alkalinity assumption (which impacts absolute CT values but not relative changes reported here).

```{r load-packages, message=FALSE, warning=FALSE}

library(tidyverse)
library(lubridate)
library(seacarb)
library(scico)

df <- read_csv(here::here("/data", "Finnmaid_mean_area_CT_2019.csv"))

df <- df %>% 
  filter(Area == "7.HGF") %>% 
  arrange(date_mean) %>% 
    mutate(day = yday(date_mean),
         day_date = as.POSIXct(strptime(paste(2000,day), format = "%Y %j",tz="GMT")),
         hour = hour(date_mean),
         week = week(date_mean),
         year = year(date_mean))

```

# Daily pCO~2~, CT and SST amplitudes

As a first step, observations are selected from all days, when the sub-region "7. HGF" was visited twice. pCO~2~, SST, CT and time differences between morning and evening visit of the sub-region are calculated. Thereafter, daytime changes per hour are calculated for pCO~2~, SST, and CT.

```{r calculate-diurnal-changes}

diurnal <-
df %>%
  add_count(year, day) %>% 
  filter(n == 2) %>% 
  group_by(year, day) %>%
  arrange(date_mean) %>%
  mutate(d_Tem = Tem_mean - lag(Tem_mean),
         d_CT = CT_mean - lag(CT_mean),
         d_pCO2 = pCO2_mean - lag(pCO2_mean),
         d_time = as.numeric(date_mean - lag(date_mean))) %>% 
  filter(!is.na(d_CT)) %>% 
  mutate(d_CT_hour = d_CT/d_time,
         d_pCO2_hour = d_pCO2/d_time,
         d_Tem_hour = d_Tem/d_time) %>% 
  ungroup()

```


```{r plot-diurnal-pCO2, fig.cap="Diurnal pCO2 changes (evening-morning) over the course of the year, pooled for 2003-2019. Color indicates corresponding SST changes."}

diurnal %>% 
ggplot(aes(day_date, d_pCO2, fill=d_Tem_hour, size=d_time))+
  geom_hline(yintercept = 0)+
  geom_point(shape = 21)+
  scale_size(name=expression(Delta~time~(hr)))+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       name=expression(Delta~SST~(K~hr^{-1})))+
  scale_x_datetime(date_labels = "%b")+
  labs(x="Month", y=expression(Delta~pCO[2]~(µatm)))+
  theme_bw()
```


```{r plot-diurnal-CT, fig.cap="Diurnal CT changes (evening-morning) over the course of the year, pooled for 2003-2019. Color indicates corresponding SST changes."}
diurnal %>% 
ggplot(aes(day_date, d_CT_hour, fill=d_Tem_hour, size=d_time))+
  geom_hline(yintercept = 0)+
  geom_point(shape = 21)+
  scale_size(name=expression(Delta~time~(hr)))+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       name=expression(Delta~SST~(K~hr^{-1})))+
  scale_x_datetime(date_labels = "%b")+
  labs(x="Month", y=expression(Delta~CT~(µmol~kg^{-1}~hr^{-1})))+
  theme_bw()


```

# Long-term weekly means of diurnal CT and SST amplitudes

In order to distill the climatological (i.e. long-term behaviour) of diurnal CT and SST amplitudes, weekly mean values of both parameters were calculated across all years.


```{r weekly-longterm-mean}

weekly <-
  diurnal %>% 
  group_by(week) %>% 
  summarise(d_CT = mean(d_CT_hour),
         d_CT_max = max(d_CT_hour),
         d_CT_min = min(d_CT_hour),
         d_Tem = mean(d_Tem_hour),
         d_Tem_max = max(d_Tem_hour),
         d_Tem_min = min(d_Tem_hour)) %>% 
  ungroup()

```


```{r plot-weekly-CT, fig.cap="Weekly mean of diurnal CT changes across all years. Color indicates corresponding SST changes."}

weekly %>% 
ggplot(aes(week, d_CT))+
  geom_hline(yintercept = 0)+
  #geom_ribbon(aes(x=week, ymax=d_CT_max, ymin = d_CT_min), alpha=0.1)+
  geom_point(aes(fill=d_Tem), shape=21, size=3)+
  geom_path()+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", name=expression(Mean~Delta~SST~(K~hr^{-1})))+
  scale_x_continuous(breaks = seq(1,60,10))+
  labs(x="Week", y=expression(Mean~Delta~CT~(µmol~kg^{-1}~hr^{-1})))+
  theme_bw()

```
# Outlook / potential open tasks

If there is an interest to quantify daytime net primary production more accurately:

* Calculate CO~2~ gas exchange occuring over the course of the day, in order to get a clean net primary production estimate. This must include a vertical mixing estimate which we already have availble for this region as an outout of high-res GETM model.

If there is an interest to derive nighttime (respiration) estimates

* Compare cummulative daytime CT changes to observed mid-term (i.e. over weeks) changes in CT, which integrate both, night- and daytime processes

If there is an interest in the controlling factors of diurnal CO~2~ variability:

* correlate observed daytime changes to environmental parameters such as mixing depth (GETM model) or radiation (remote sensing)




